White Dwarf Mergers on Adaptive Meshes\\ II. Collisions

### 2017-07-09 papers/wdmerger_II/paper.tex
Describe the forced post-step regrid algorithm

We agree with \citet{kushnir:2013} that solving this numerical
instability is crucial to avoiding unphysical detonations.
A simulation that does not solve this problem will not obtain
the correct amount and will not converge properly with resolution.
This may justify the addition of many AMR levels to the domain
if a correct evaluation of the burning phase is desired.
Even if one cannot afford the full resolution required by this
AMR criterion, and chooses to limit the number of levels to some
predetermined maximum based on a constraint of computing time,
the added resolution will still go to the most-needed places.
However, a significant drawback to this approach is that it
does not turn on until after an unstable burning region has
been generated, so this AMR criterion cannot help in the case
where a spurious detonation begins in a single timestep on the coarse
grid, though it will kick in immediately after that timestep
to resolve the regions where the detonation has occurred.
A possible remedy to be explored in future work is to add
refinement pre-emptively in regions where the ratio $\tau_s / \tau_e$
approaches $f_{s}$ but does not yet exceed it. Another
possibility for this case would be to track how many levels
would have been needed to prevent the numerical detonation,
based on the stability criterion, and reject the timestep and
retry it from the starting state using that many levels of refinement.
As mentioned in \autoref{sec:hydrocoupling}, the ability to reject
a timestep and retry it is functionality that
we have recently added to \castro, so we may try this in future work.

>>>>

We agree with \citet{kushnir:2013} that solving this numerical
instability is crucial to avoiding unphysical detonations.
A simulation that does not solve this problem will not obtain
the correct amount and will not converge properly with resolution.
This may justify the addition of many AMR levels to the domain
if a correct evaluation of the burning phase is desired.
Even if one cannot afford the full resolution required by this
AMR criterion, and chooses to limit the number of levels to some
predetermined maximum based on a constraint of computing time,
the added resolution will still go to the most-needed places.

A potential limitation of this approach is that it does not turn on
until \textit{after} an unstable burning region has been generated,
because we generally only perform a regridding and level generation
step at times when AMR levels are synchronized, such as at the beginning
of the timestep on the coarsest level. So this AMR criterion cannot
help in the case where a spurious detonation begins in a single timestep
on the coarse grid, though it will kick in immediately after that
timestep to resolve the regions where the detonation has occurred.
This situation is common in our white dwarf collision simulations.
We have developed a remedy to this issue. At the \textit{end} of every step
on every level, we check whether the stability criterion has been violated.
If so, we recognize that it is not too late to solve the problem -- we can
still add levels of refinement in those places. We then perform the new level
generation to cover specifically those places that violated the stability criterion.
To ensure consistency with our algorithm, we use the \textit{old}-time values
of the coarse fluid state to interpolate into the new fine level, even
though it is the \textit{new}-time values that were used to guide the level
generation. Then our algorithm evolves the timestep on the new level as normal.
Because our algorithm is recursive, this can continue to generate
as many new levels as are needed to satisfy the stability criterion.
The upshot is that as long as the simulation can afford enough levels to
satisfy the stability criterion, there never needs to be even a single timestep
that violates the criterion.
